---
title: "R Notebook"
output: html_notebook
---


Align 

```{r}
library(tidyverse)
library(survival)
library(survminer)
```

```{r}
# Load data 
setwd("/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/MSK_survival") 

clin_pt  <- read.delim("data_clinical_patient.txt", skip = 4, check.names = FALSE)
clin_smp <- read.delim("data_clinical_sample.txt",  skip = 4, check.names = FALSE)
cnv_raw  <- read.delim("all_cnv_table.txt", check.names = FALSE) %>%
  rownames_to_column(var = "SAMPLE_ID")

arm_cols      <- grep("^Arm",           names(cnv_raw), value = TRUE)
cytoband_cols <- grep("^[0-9XY]+[pq]", names(cnv_raw), value = TRUE)
```

```{r}
# Compute AS. Use AS as a covariate 
aneuploidy_cytoband <- cnv_raw %>%
  select(SAMPLE_ID, all_of(cytoband_cols)) %>%
  mutate(AneuploidyScore = rowSums(abs(across(all_of(cytoband_cols))), na.rm = TRUE)) %>%
  select(SAMPLE_ID, AneuploidyScore)

```

```{r}
#Build df
sample_info_extended <- clin_smp %>%
  select(SAMPLE_ID, PATIENT_ID, CANCER_TYPE, ONCOTREE_CODE, TUMOR_PURITY) %>%
  left_join(aneuploidy_cytoband, by = "SAMPLE_ID")

# Categorize CNVs
cnv_cat_cytoband <- cnv_raw %>%
  mutate(across(all_of(cytoband_cols), ~ case_when(
    .x >  0.2 ~ "Gain",
    .x < -0.2 ~ "Loss",
    TRUE       ~ "None"
  )))

merged_sample_cytoband <- merge(sample_info_extended, cnv_cat_cytoband, by = "SAMPLE_ID")

df_cyto <- merge(
  clin_pt[, c("PATIENT_ID", "OS_MONTHS", "OS_STATUS", "DRUG_TYPE")],
  merged_sample_cytoband,
  by = "PATIENT_ID"
) %>%
  mutate(OS = if_else(str_detect(OS_STATUS, "LIVING"), 0, 1))

head(df_cyto)
```

```{r}
# Cox regression loop
cancer_types_to_plot <- c(
  "Head and Neck Cancer", "Melanoma",
  "Non-Small Cell Lung Cancer", "Renal Cell Carcinoma"
)

results_cytoband <- data.frame()

for (tumor in cancer_types_to_plot) {
  tumor_data <- df_cyto %>%
    filter(CANCER_TYPE == tumor, DRUG_TYPE == "PD-1/PDL-1")

  for (band in cytoband_cols) {
    for (comparison in c("Gain", "Loss")) {
      subset <- tumor_data %>%
        filter(!!sym(band) %in% c("None", comparison)) %>%
        mutate(CNV = factor(!!sym(band), levels = c(comparison, "None"))) %>%
        filter(!is.na(TUMOR_PURITY), !is.na(AneuploidyScore))

      if (nrow(subset) >= 10 && length(unique(subset$CNV)) == 2) {
        tryCatch({
          fit <- coxph(Surv(OS_MONTHS, OS) ~ CNV + TUMOR_PURITY + AneuploidyScore,
                       data = subset)
          cnv_term <- grep("^CNV", names(coef(fit)), value = TRUE)
          if (length(cnv_term) != 1) next

          hr    <- exp(coef(fit)[cnv_term])
          ci    <- round(exp(confint(fit)[cnv_term, ]), 3)
          p     <- summary(fit)$coefficients[cnv_term, 5]

          results_cytoband <- rbind(results_cytoband, data.frame(
            CANCER_TYPE = tumor,
            Cytoband    = band,
            Comparison  = comparison,
            HR          = round(hr, 3),
            log2HR      = log2(hr),
            sign_logP   = -log10(p) * sign(log2(hr)),
            CI          = paste(ci[1], ci[2], sep = " - "),
            P           = signif(p, 3),
            N           = nrow(subset)
          ))
        }, error = function(e) {
          message(paste("Skipped:", tumor, band, comparison, "-", e$message))
        })
      }
    }
  }
}

cat("Total results:", nrow(results_cytoband), "\n")
```


```{r}
# Find shared cytobands across ALL tumor types  
shared_cytobands <- results_cytoband %>%
  group_by(Cytoband) %>%
  summarise(n_tumors = n_distinct(CANCER_TYPE), .groups = "drop") %>%
  filter(n_tumors == length(unique(results_cytoband$CANCER_TYPE))) %>%
  pull(Cytoband)

cat("Shared cytobands:", length(shared_cytobands), "\n")

# Filter to shared cytobands only
results_cytoband_shared <- results_cytoband %>%
  filter(Cytoband %in% shared_cytobands) %>%
  mutate(
    Event      = ifelse(Comparison == "Gain", "Gain", "Loss"),
    ArmLabel   = str_extract(Cytoband, "^[0-9XY]+[pq]"),
    chr_num    = as.numeric(str_extract(Cytoband, "^[0-9]+")),
    chr_num    = ifelse(is.na(chr_num),
                        ifelse(str_detect(Cytoband, "^X"), 23, 24),
                        chr_num),
    arm_suffix = str_extract(Cytoband, "[pq]")
  )

# Genomic order
cyto_levels_shared <- results_cytoband_shared %>%
  arrange(chr_num, arm_suffix, Cytoband) %>%
  pull(Cytoband) %>%
  unique()

results_cytoband_shared$Cytoband <- factor(
  results_cytoband_shared$Cytoband,
  levels = cyto_levels_shared
)

# One label per arm
axis_labels_shared <- results_cytoband_shared %>%
  group_by(ArmLabel) %>%
  summarise(Cytoband = Cytoband[ceiling(n() / 2)], .groups = "drop")
```


```{r}
# Plot 
#dir.create("logHR_plots_updated", showWarnings = FALSE)

for (tumor in unique(results_cytoband_shared$CANCER_TYPE)) {
  plot_data <- results_cytoband_shared %>% filter(CANCER_TYPE == tumor)

  p <- ggplot(plot_data, aes(x = Cytoband, y = sign_logP, fill = Event)) +
    geom_bar(stat = "identity", width = 1) +

    # Chromosome boundary lines (vertical)
    geom_vline(xintercept = which(levels(plot_data$Cytoband) %in%
                 (plot_data %>%
                    group_by(chr_num) %>%
                    summarise(first = Cytoband[1], .groups = "drop") %>%
                    pull(first))),
               color = "grey60", linewidth = 0.3, linetype = "dashed") +

    # Significance lines
    geom_hline(yintercept =  -log10(0.05), linetype = "dotted", color = "grey60", linewidth = 0.3) +
    geom_hline(yintercept =   log10(0.05), linetype = "dotted", color = "grey60", linewidth = 0.3) +
    geom_hline(yintercept =  -log10(0.01), linetype = "dotted", color = "grey40", linewidth = 0.3) +
    geom_hline(yintercept =   log10(0.01), linetype = "dotted", color = "grey40", linewidth = 0.3) +
    geom_hline(yintercept = 0, linetype = "solid", linewidth = 0.4, color = "black") +

    annotate("text", x = Inf, y = -log10(0.05), label = "p<0.05",
             hjust = 1.1, size = 2.8, color = "grey40") +
    annotate("text", x = Inf, y =  log10(0.05), label = "p<0.05",
             hjust = 1.1, size = 2.8, color = "grey40") +
    annotate("text", x = Inf, y = -log10(0.01), label = "p<0.01",
             hjust = 1.1, size = 2.8, color = "grey20") +
    annotate("text", x = Inf, y =  log10(0.01), label = "p<0.01",
             hjust = 1.1, size = 2.8, color = "grey20") +

    scale_fill_manual(values = c("Gain" = "#E8437A", "Loss" = "#5BC4E5")) +
    scale_x_discrete(
      breaks = axis_labels_shared$Cytoband,
      labels = axis_labels_shared$ArmLabel
    ) +
    scale_y_continuous(limits = c(-2, 4), breaks = c(-2, 0, 2, 4)) +
    theme_bw()+
    labs(
      title = paste0(tumor, " – Association with Patient Survival"),
      x     = "Chromosome",
      y     = "Signed –log10(p)",
      fill  = "Event"
    ) +
    theme(
      axis.text.x        = element_text(angle = 90, size = 6),
      plot.title         = element_text(face = "bold", size = 12),
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "white"),
      plot.background  = element_rect(fill = "white"),
      panel.border     = element_blank(),          
      axis.line        = element_line(color = "black", linewidth = 0.4) 
    )

  ggsave(
    filename = paste0("logHR_plots_updated/", gsub(" ", "_", tumor), ".pdf"),
    plot = p, width = 10, height = 4
  )
}
```


```{r}
# Check factor levels in new file for one example
subset_test <- df_cyto %>%
  filter(CANCER_TYPE == "Melanoma", DRUG_TYPE == "PD-1/PDL-1") %>%
  filter(!!sym("1q21.1") %in% c("None", "Gain")) %>%
  mutate(CNV = factor(!!sym("1q21.1"), levels = c("Gain", "None")))

cat("Factor levels:", levels(subset_test$CNV), "\n")
# Should be "Gain" "None" — Gain is reference
# HR > 1 means Gain has worse survival than None
```

```{r}
# Check a specific cytoband manually
subset_test <- df_cyto %>%
  filter(CANCER_TYPE == "Melanoma", DRUG_TYPE == "PD-1/PDL-1") %>%
  filter(`1q21.1` %in% c("None", "Gain")) %>%
  mutate(CNV = factor(`1q21.1`, levels = c("Gain", "None"))) %>%
  filter(!is.na(TUMOR_PURITY), !is.na(AneuploidyScore))

fit <- coxph(Surv(OS_MONTHS, OS) ~ CNV + TUMOR_PURITY + AneuploidyScore, data = subset_test)
cnv_term <- grep("^CNV", names(coef(fit)), value = TRUE)

hr <- exp(coef(fit)[cnv_term])
p  <- summary(fit)$coefficients[cnv_term, 5]

cat("HR:", hr, "\n")
cat("log2HR:", log2(hr), "\n")
cat("sign:", sign(log2(hr)), "\n")
cat("sign_logP:", -log10(p) * sign(log2(hr)), "\n")
```

# Figure 5 Align BLCA 
## arm level 
```{r}
# ── Build everything from scratch ─────────────────────────────────────────────
setwd("/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/MSK_survival")

clin_pt  <- read.delim("data_clinical_patient.txt", skip = 4, check.names = FALSE)
clin_smp <- read.delim("data_clinical_sample.txt",  skip = 4, check.names = FALSE)
cnv_raw  <- read.delim("all_cnv_table.txt", check.names = FALSE) %>%
  rownames_to_column(var = "SAMPLE_ID")

arm_cols      <- grep("^Arm",           names(cnv_raw), value = TRUE)
cytoband_cols <- grep("^[0-9XY]+[pq]", names(cnv_raw), value = TRUE)

cat("Arm11p in arm_cols:", "Arm11p" %in% arm_cols, "\n")
cat("Total arm cols:", length(arm_cols), "\n")
```

```{r}
# Categorize CNVs for BOTH arm and cytoband columns
cnv_cat <- cnv_raw %>%
  mutate(across(all_of(c(arm_cols, cytoband_cols)), ~ case_when(
    .x >  0.2 ~ "Gain",
    .x < -0.2 ~ "Loss",
    TRUE       ~ "None"
  )))
```

```{r}
# Build df_arm 
sample_info <- clin_smp %>%
  select(SAMPLE_ID, PATIENT_ID, CANCER_TYPE, ONCOTREE_CODE)

merged_sample_arm <- sample_info %>%
  left_join(cnv_cat, by = "SAMPLE_ID")    # CHANGED: cnv_cat_arm → cnv_cat

df_arm <- clin_pt %>%
  select(PATIENT_ID, OS_MONTHS, OS_STATUS, DRUG_TYPE) %>%
  left_join(merged_sample_arm, by = "PATIENT_ID") %>%
  mutate(OS = if_else(str_detect(OS_STATUS, "LIVING"), 0, 1))

cat("Rows in df_arm:", nrow(df_arm), "\n")
cat("DRUG_TYPE values:", unique(df_arm$DRUG_TYPE), "\n")
cat("Arm11p column exists:", "Arm11p" %in% colnames(df_arm), "\n")
```

```{r}
# Filter to BLCA + PD-1/PDL-1 in BLCA
BLCA_data <- df_arm %>%
  filter(CANCER_TYPE == "Bladder Cancer", DRUG_TYPE == "PD-1/PDL-1")

cat("Rows in tumor_data:", nrow(BLCA_data), "\n")
table(BLCA_data$Arm11p)
```
```{r}
# ── Arm-level Cox regression for BLCA across all arms ────────────────────────
results_blca_armlevel <- data.frame()

for (arm in arm_cols) {
  for (comparison in c("Gain", "Loss")) {
    subset <- BLCA_data %>%
      filter(!!sym(arm) %in% c("None", comparison)) %>%
      mutate(CNV = factor(!!sym(arm), levels = c("None", comparison)))

    if (nrow(subset) >= 10 && length(unique(subset$CNV)) == 2) {
      tryCatch({
        fit <- coxph(Surv(OS_MONTHS, OS) ~ CNV, data = subset)
        cnv_term <- grep("^CNV", names(coef(fit)), value = TRUE)
        if (length(cnv_term) != 1) next

        hr  <- exp(coef(fit)[cnv_term])
        p   <- summary(fit)$coefficients[cnv_term, 5]

        results_blca_armlevel <- rbind(results_blca_armlevel, data.frame(
          CANCER_TYPE = "Bladder Cancer",
          Arm         = arm,
          Comparison  = comparison,
          HR          = round(hr, 3),
          log2HR      = log2(hr),
          sign_logP   = -log10(p) * sign(log2(hr)),
          P           = signif(p, 3),
          N           = nrow(subset)
        ))
      }, error = function(e) {
        message(paste("Skipped:", arm, comparison, "-", e$message))
      })
    }
  }
}

cat("Arm-level results:", nrow(results_blca_armlevel), "\n")
results_blca_armlevel %>% filter(Arm == "Arm11p")
```

```{r}
# ── Plot BLCA survival ────────────────────────────────────────────────────────
results_blca_armlevel <- results_blca_armlevel %>%
  mutate(
    Event      = ifelse(Comparison == "Gain", "Gain", "Loss"),
    ArmName    = gsub("Arm", "", Arm),
    chr_num    = as.numeric(str_extract(ArmName, "^[0-9]+")),
    chr_num    = ifelse(is.na(chr_num),
                        ifelse(str_detect(ArmName, "^X"), 23, 24),
                        chr_num),
    arm_suffix = str_extract(ArmName, "[pq]")
  )

arm_levels <- results_blca_armlevel %>%
  arrange(chr_num, arm_suffix) %>%
  pull(ArmName) %>%
  unique()

results_blca_armlevel$ArmName <- factor(results_blca_armlevel$ArmName, levels = arm_levels)

p_blca_arm <- ggplot(results_blca_armlevel, aes(x = ArmName, y = sign_logP, fill = Event)) +
  geom_bar(stat = "identity", width = 1) +
  geom_hline(yintercept =  -log10(0.05), linetype = "dotted", color = "grey60") +
  geom_hline(yintercept =   log10(0.05), linetype = "dotted", color = "grey60") +
  geom_hline(yintercept =  -log10(0.01), linetype = "dotted", color = "grey40") +
  geom_hline(yintercept =   log10(0.01), linetype = "dotted", color = "grey40") +
  geom_hline(yintercept = 0, linetype = "solid") +
  scale_fill_manual(values = c("Gain" = "#E8437A", "Loss" = "#5BC4E5")) +
  labs(
    title = "BLCA (bladder) – Arm-level Association with Patients' Survival",
    x     = "Chromosome Arm",
    y     = "Signed –log10(p)",
    fill  = "Event"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, size = 7),
    plot.title  = element_text(face = "bold", size = 12)
  )

print(p_blca_arm)

ggsave("/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/MSK_survival/logHR_plots_updated/BLCA_survival_armlevel.pdf",
       plot = p_blca_arm, width = 12, height = 4)
```

## cytoband level 
```{r}
is_cnv_blca <- read.csv(
  "/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/TCGA/Resubmission_SCNAs/IS_CNV_cytolevel.csv",
  check.names = FALSE
)

# Get the cytobands from IS_CNV file
is_cnv_cytobands <- unique(is_cnv_blca$Cytoband)
cat("Cytobands in IS_CNV:", length(is_cnv_cytobands), "\n")
```

```{r}
cat("Cytoband cols found:", length(cytoband_cols), "\n")
cat("11p15.5 in cytoband_cols:", "11p15.5" %in% cytoband_cols, "\n")
```

```{r}
# Categorize CYTOBAND columns only
cnv_cat_cytoband <- cnv_raw %>%
  mutate(across(all_of(cytoband_cols), ~ case_when(
    .x >  0.2 ~ "Gain",
    .x < -0.2 ~ "Loss",
    TRUE       ~ "None"
  )))

cat("Class of 11p15.5 after categorization:", class(cnv_cat_cytoband$`11p15.5`), "\n")
table(cnv_cat_cytoband$`11p15.5`)
```
```{r}
# Build df_cyto
sample_info <- clin_smp %>%
  select(SAMPLE_ID, PATIENT_ID, CANCER_TYPE, ONCOTREE_CODE, TUMOR_PURITY)

merged_sample_cytoband <- sample_info %>%
  left_join(cnv_cat_cytoband, by = "SAMPLE_ID")

df_cyto <- clin_pt %>%
  select(PATIENT_ID, OS_MONTHS, OS_STATUS, DRUG_TYPE) %>%
  left_join(merged_sample_cytoband, by = "PATIENT_ID") %>%
  mutate(OS = if_else(str_detect(OS_STATUS, "LIVING"), 0, 1))

cat("Rows in df_cyto:", nrow(df_cyto), "\n")
cat("Class of 11p15.5 in df_cyto:", class(df_cyto$`11p15.5`), "\n")
table(df_cyto$`11p15.5`)
```


```{r}
# Run Cox regression for BLCA  
# First build df_cyto as before (same setup chunks)
# Then filter to only cytobands present in IS_CNV file

results_blca <- data.frame()

tumor_data <- df_cyto %>%
  filter(CANCER_TYPE == "Bladder Cancer", DRUG_TYPE == "PD-1/PDL-1")

# Only run Cox for cytobands that exist in IS_CNV file
shared_bands <- intersect(cytoband_cols, is_cnv_cytobands)
cat("Shared cytobands between MSK and IS_CNV:", length(shared_bands), "\n")

for (band in shared_bands) {
  for (comparison in c("Gain", "Loss")) {
    subset <- tumor_data %>%
      filter(!!sym(band) %in% c("None", comparison)) %>%
      mutate(CNV = factor(!!sym(band), levels = c("None", comparison)))  
      #filter(!is.na(TUMOR_PURITY), !is.na(AneuploidyScore))

    if (nrow(subset) >= 10 && length(unique(subset$CNV)) == 2) {
      tryCatch({
        fit <- coxph(Surv(OS_MONTHS, OS) ~ CNV,
                     data = subset)
        cnv_term <- grep("^CNV", names(coef(fit)), value = TRUE)
        if (length(cnv_term) != 1) next

        hr  <- exp(coef(fit)[cnv_term])
        p   <- summary(fit)$coefficients[cnv_term, 5]

        results_blca <- rbind(results_blca, data.frame(
          CANCER_TYPE = "Bladder Cancer",
          Cytoband    = band,
          Comparison  = comparison,
          HR          = round(hr, 3),
          log2HR      = log2(hr),
          sign_logP   = -log10(p) * sign(log2(hr)),
          P           = signif(p, 3),
          N           = nrow(subset)
        ))
      }, error = function(e) {
        message(paste("Skipped:", band, comparison, "-", e$message))
      })
    }
  }
}

cat("BLCA survival results:", nrow(results_blca), "\n")
```


```{r}
# Plot WITHOUT filling missing cytobands - using results_blca BEFORE the fill-in step
# (only the 1595 rows you originally got from the Cox loop)

results_blca_nofill <- results_blca_complete %>%
  filter(N > 0)  # keep only cytobands that actually had a Cox model run

cat("Rows without fill-in:", nrow(results_blca_nofill), "\n")

# Order cytobands (only the ones that exist)
results_blca_nofill <- results_blca_nofill %>%
  mutate(
    ArmLabel   = str_extract(Cytoband, "^[0-9XY]+[pq]"),
    chr_num    = as.numeric(str_extract(Cytoband, "^[0-9]+")),
    chr_num    = ifelse(is.na(chr_num),
                        ifelse(str_detect(Cytoband, "^X"), 23, 24),
                        chr_num),
    arm_suffix = str_extract(Cytoband, "[pq]")
  )

cyto_levels_nofill <- results_blca_nofill %>%
  arrange(chr_num, arm_suffix, Cytoband) %>%
  pull(Cytoband) %>%
  unique()

results_blca_nofill$Cytoband <- factor(results_blca_nofill$Cytoband, levels = cyto_levels_nofill)

axis_labels_nofill <- results_blca_nofill %>%
  group_by(ArmLabel) %>%
  summarise(Cytoband = Cytoband[ceiling(n() / 2)], .groups = "drop")

# Plot
p_blca_nofill <- ggplot(results_blca_nofill, aes(x = Cytoband, y = sign_logP, fill = Event, group = Event)) +
  geom_area(alpha = 0.8) +
  geom_hline(yintercept = 0, linetype = "solid", linewidth = 0.4, color = "black") +
  scale_fill_manual(values  = c("Gain" = "#E8437A", "Loss" = "#5BC4E5")) +
  scale_x_discrete(
    breaks = axis_labels_nofill$Cytoband,
    labels = axis_labels_nofill$ArmLabel
  ) +
  labs(
    title = "BLCA – Survival (NO fill-in, real data only)",
    x = "Chromosome", y = "Signed –log10(p)", fill = "Event"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, size = 6))

print(p_blca_nofill)
```




## align 
```{r}
# ── Fill in missing cytobands with 0 so all 805 shared bands are present ──────
all_combinations <- expand.grid(
  Cytoband   = shared_bands,
  Comparison = c("Gain", "Loss"),
  stringsAsFactors = FALSE
)

results_blca_complete <- all_combinations %>%
  left_join(results_blca, by = c("Cytoband", "Comparison")) %>%
  mutate(
    CANCER_TYPE = "Bladder Cancer",
    HR          = ifelse(is.na(HR),       1, HR),
    log2HR      = ifelse(is.na(log2HR),   0, log2HR),
    sign_logP   = ifelse(is.na(sign_logP),0, sign_logP),
    P           = ifelse(is.na(P),        1, P),
    N           = ifelse(is.na(N),        0, N),
    Event       = ifelse(Comparison == "Gain", "Gain", "Loss")
  )

# Replace results_blca with complete version
results_blca <- results_blca_complete

cat("Complete BLCA results:", nrow(results_blca), "\n")
cat("Expected:", length(shared_bands) * 2, "\n")
```

```{r}
# ── Order cytobands to match IS_CNV file order ────────────────────────────────
results_blca <- results_blca %>%
  mutate(
    Event      = ifelse(Comparison == "Gain", "Gain", "Loss"),
    ArmLabel   = str_extract(Cytoband, "^[0-9XY]+[pq]"),
    chr_num    = as.numeric(str_extract(Cytoband, "^[0-9]+")),
    chr_num    = ifelse(is.na(chr_num),
                        ifelse(str_detect(Cytoband, "^X"), 23, 24),
                        chr_num),
    arm_suffix = str_extract(Cytoband, "[pq]")
  )

# Genomic order matching IS_CNV
cyto_levels_blca <- results_blca %>%
  arrange(chr_num, arm_suffix, Cytoband) %>%
  pull(Cytoband) %>%
  unique()

results_blca$Cytoband <- factor(results_blca$Cytoband, levels = cyto_levels_blca)

axis_labels_blca <- results_blca %>%
  group_by(ArmLabel) %>%
  summarise(Cytoband = Cytoband[ceiling(n() / 2)], .groups = "drop")
```

```{r}
# ── Plot BLCA survival ────────────────────────────────────────────────────────
p_blca <- ggplot(results_blca, aes(x = Cytoband, y = sign_logP, fill = Event, group = Event)) +
  geom_area(alpha = 0.8) +    # filled area
  #geom_line(aes(color = Event), linewidth = 0.3) + # outline
  geom_vline(xintercept = which(levels(results_blca$Cytoband) %in%
               (results_blca %>%
                  group_by(chr_num) %>%
                  summarise(first = Cytoband[1], .groups = "drop") %>%
                  pull(first))),
             color = "grey60", linewidth = 0.3, linetype = "dashed") +
  geom_hline(yintercept =  -log10(0.05), linetype = "dotted", color = "grey60", linewidth = 0.3) +
  geom_hline(yintercept =   log10(0.05), linetype = "dotted", color = "grey60", linewidth = 0.3) +
  geom_hline(yintercept =  -log10(0.01), linetype = "dotted", color = "grey40", linewidth = 0.3) +
  geom_hline(yintercept =   log10(0.01), linetype = "dotted", color = "grey40", linewidth = 0.3) +
  geom_hline(yintercept = 0, linetype = "solid", linewidth = 0.4, color = "black") +
  scale_fill_manual(values  = c("Gain" = "#E8437A", "Loss" = "#5BC4E5")) +
  #scale_color_manual(values = c("Gain" = "#E8437A", "Loss" = "#5BC4E5")) +
  scale_x_discrete(
    breaks = axis_labels_blca$Cytoband,
    labels = axis_labels_blca$ArmLabel
  ) +
  scale_y_continuous(limits = c(-2, 4), breaks = c(-2, 0, 2, 4)) +
  labs(
    title = "BLCA (bladder) – Association with Patients' Survival – anti PD1/PD-L1",
    x     = "Chromosome",
    y     = "Signed –log10(p)",
    fill  = "Event"
  ) +
  guides(color = "none") +   # hide duplicate color legend
  theme_bw() +
  theme(
    axis.text.x      = element_text(angle = 90, size = 6),
    plot.title       = element_text(face = "bold", size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white"),
    plot.background  = element_rect(fill = "white"),
    panel.border     = element_blank(),
    axis.line        = element_line(color = "black", linewidth = 0.4)
  )

ggsave("/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/MSK_survival/logHR_plots_updated/BLCA_survival_aligned.pdf",
       plot = p_blca, width = 12, height = 4)

p_blca 
```


```{r}
cytobands_11p <- grep("^11p", cytoband_cols, value = TRUE)
cat("11p cytobands:", cytobands_11p, "\n\n")

results_11p_check <- data.frame()

for (band in cytobands_11p) {
  for (comparison in c("Gain", "Loss")) {

    subset_unadj <- tumor_data %>%
      filter(!!sym(band) %in% c("None", comparison)) %>%
      mutate(CNV = factor(!!sym(band), levels = c("None", comparison)))

    subset_adj <- subset_unadj %>%
      filter(!is.na(TUMOR_PURITY), !is.na(AneuploidyScore))

    if (nrow(subset_unadj) >= 10 && length(unique(subset_unadj$CNV)) == 2) {

      p_unadj <- NA
      p_adj   <- NA

      # Unadjusted
      tryCatch({
        fit_unadj <- coxph(Surv(OS_MONTHS, OS) ~ CNV, data = subset_unadj)
        cnv_term  <- grep("^CNV", names(coef(fit_unadj)), value = TRUE)
        if (length(cnv_term) == 1 && is.finite(coef(fit_unadj)[cnv_term])) {
          p_unadj <- summary(fit_unadj)$coefficients[cnv_term, "Pr(>|z|)"]
        }
      }, error = function(e) {
        message(paste("Unadjusted failed:", band, comparison, "-", e$message))
      })

      # Adjusted
      if (nrow(subset_adj) >= 10 && length(unique(subset_adj$CNV)) == 2) {
        tryCatch({
          fit_adj  <- coxph(Surv(OS_MONTHS, OS) ~ CNV + TUMOR_PURITY + AneuploidyScore, data = subset_adj)
          cnv_term <- grep("^CNV", names(coef(fit_adj)), value = TRUE)
          if (length(cnv_term) == 1 && is.finite(coef(fit_adj)[cnv_term])) {
            p_adj <- summary(fit_adj)$coefficients[cnv_term, "Pr(>|z|)"]
          }
        }, error = function(e) {
          message(paste("Adjusted failed:", band, comparison, "-", e$message))
        })
      }

      results_11p_check <- rbind(results_11p_check, data.frame(
        Cytoband      = band,
        Comparison    = comparison,
        N_unadj       = nrow(subset_unadj),
        N_adj         = nrow(subset_adj),
        P_unadjusted  = signif(p_unadj, 3),
        P_adjusted    = signif(p_adj, 3)
      ))
    }
  }
}

print(results_11p_check)
```




# threshold 0.3
```{r}
library(gtools)  # for mixedsort()

# ── Rebuild df with 0.3 threshold (matching colleague's code) ────────────────
setwd("/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/MSK_survival")

clin_pt  <- read.delim("data_clinical_patient.txt", skip = 4, check.names = FALSE)
clin_smp <- read.delim("data_clinical_sample.txt", skip = 4, check.names = FALSE)
cnv_raw  <- read.delim("all_cnv_table.txt", check.names = FALSE, quote = "\"") %>%
  tibble::rownames_to_column("SAMPLE_ID")

clin_smp$SAMPLE_ID <- as.character(clin_smp$SAMPLE_ID)
cnv_raw$SAMPLE_ID  <- as.character(cnv_raw$SAMPLE_ID)

cytoband_cols <- grep("^[0-9XY]+[pq]", names(cnv_raw), value = TRUE)
arm_cols      <- grep("^Arm", names(cnv_raw), value = TRUE)

# Categorize at 0.3 threshold (NOT 0.2)
cnv_cat <- cnv_raw %>%
  mutate(across(all_of(c(cytoband_cols, arm_cols)), ~ case_when(
    .x >  0.3 ~ "Gain",
    .x < -0.3 ~ "Loss",
    TRUE      ~ "None"
  )))

sample_info <- clin_smp %>%
  select(SAMPLE_ID, PATIENT_ID, CANCER_TYPE, ONCOTREE_CODE, SAMPLE_TYPE)

merged_sample <- sample_info %>%
  left_join(cnv_cat, by = "SAMPLE_ID")

df <- clin_pt %>%
  select(PATIENT_ID, OS_MONTHS, OS_STATUS, SEX, DRUG_TYPE) %>%
  left_join(merged_sample, by = "PATIENT_ID") %>%
  mutate(OS = if_else(str_detect(OS_STATUS, "LIVING"), 0, 1))

cat("Total patients:", length(unique(df$PATIENT_ID)), "\n")
```

```{r}
# Run Cox regression for BLCA cytobands  
results_cytoband <- data.frame()

tumor_data <- df %>% filter(CANCER_TYPE == "Bladder Cancer", DRUG_TYPE == "PD-1/PDL-1")

for (band in cytoband_cols) {
  for (comparison in c("Gain", "Loss")) {
    subset <- tumor_data %>%
      filter(!!sym(band) %in% c("None", comparison)) %>%
      mutate(CNV = factor(!!sym(band), levels = c("None", comparison)))

    if (nrow(subset) >= 10 && length(unique(subset$CNV)) == 2) {
      tryCatch({
        fit <- coxph(Surv(OS_MONTHS, OS) ~ CNV, data = subset)
        hr  <- exp(coef(fit))
        p   <- summary(fit)$coefficients[5]

        results_cytoband <- rbind(results_cytoband, data.frame(
          CANCER_TYPE = "Bladder Cancer",
          Cytoband    = band,
          Comparison  = comparison,
          HR          = round(hr, 3),
          CI          = paste(round(exp(confint(fit)), 3)[1],
                              round(exp(confint(fit)), 3)[2], sep = " - "),
          P           = signif(p, 3),
          log2HR      = log2(hr),
          Event       = comparison,
          N           = nrow(subset),
          sign_logP   = -log10(p) * sign(log2(hr))
        ))
      }, error = function(e) {
        # Still add the cytoband with 0 if Cox fails
        results_cytoband <<- rbind(results_cytoband, data.frame(
          CANCER_TYPE = "Bladder Cancer",
          Cytoband    = band,
          Comparison  = comparison,
          HR          = 1, CI = "0", P = 1,
          log2HR      = 0, Event = comparison,
          N           = nrow(subset), sign_logP = 0
        ))
      })
    } else {
      # ADD cytobands that fail the filter with sign_logP = 0
      results_cytoband <- rbind(results_cytoband, data.frame(
        CANCER_TYPE = "Bladder Cancer",
        Cytoband    = band,
        Comparison  = comparison,
        HR          = 1, CI = "0", P = 1,
        log2HR      = 0, Event = comparison,
        N           = nrow(subset), sign_logP = 0
      ))
    }
  }
}

cat("BLCA results:", nrow(results_cytoband), "\n")
cat("Expected:", length(cytoband_cols) * 2, "\n")
# Add signed -log10(p)
results_cytoband <- results_cytoband %>%
  mutate(sign_logP = -log10(P) * sign(log2HR))

cat("BLCA results:", nrow(results_cytoband), "\n")
```

```{r}
# Plot
out_dir <- "/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/MSK_survival/logHR_plots_updated"
if (!dir.exists(out_dir)) dir.create(out_dir)

plot_cnv_area <- function(df, tumour, signif_cut = 1.3) {

  chr_levels <- c(as.character(1:22), "X", "Y")

  pd <- df %>%
    filter(CANCER_TYPE == tumour) %>%
    mutate(Arm        = str_extract(Cytoband, "^[0-9XY]+[pq]"),
           Chromosome = str_extract(Arm, "^[0-9XY]+"),
           Chromosome = factor(Chromosome, levels = chr_levels)) %>%
    arrange(Chromosome, Cytoband) %>%
    mutate(Cytoband = factor(Cytoband, levels = mixedsort(unique(Cytoband))),
           idx      = row_number())

  centromere_idx <- pd %>% filter(str_detect(Arm, "q")) %>%
    group_by(Chromosome) %>% slice_head(n = 1) %>% pull(idx)

  chr_last_idx <- pd %>% group_by(Chromosome) %>%
    slice_tail(n = 1) %>% pull(idx)

  chr_mid_idx <- pd %>% group_by(Chromosome) %>%
    summarise(idx = floor(mean(range(idx))), .groups = "drop")

  ggplot(pd, aes(idx, sign_logP, fill = Event)) +
    geom_area(position = "identity", alpha = 0.75) +
    geom_hline(yintercept = 0, linewidth = .25) +
    geom_hline(yintercept = c(-signif_cut, signif_cut),
               colour = "grey40", linetype = "dashed", linewidth = .25) +
    geom_vline(xintercept = chr_last_idx + 0.5, colour = "grey30", linetype = "dashed",linewidth = .3) +
    geom_vline(xintercept = centromere_idx, colour = "grey",
               linetype = "dashed", linewidth = .2) +
    scale_fill_manual(values = c("Gain" = "#E8437A", "Loss" = "#5BC4E5")) +
    scale_x_continuous(breaks = chr_mid_idx$idx,
                       labels = chr_mid_idx$Chromosome,
                       expand = c(0, 0)) +
    labs(title = paste("Signed -log10(p) CNV Survival –", tumour),
         x = "chromosome",
         y = "signed -log(p)",
         fill = "Event") +
    theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text.x      = element_text(size = 6),
          plot.title       = element_text(face = "bold", size = 12))
}

p_blca <- plot_cnv_area(results_cytoband, "Bladder Cancer")
print(p_blca)

ggsave(file.path(out_dir, "Bladder_Cancer_0.3_no_covariate.pdf"),
       p_blca, width = 10, height = 5, useDingbats = FALSE)
```

## use padj 
```{r}
# Apply padj 
results_cytoband_padj <- results_cytoband %>%
  mutate(
    sign_logP = -log10(p.adjust(P, method = "BH")) * sign(log2HR)
  )

cat("rows:", nrow(results_cytoband_padj), "\n")
```

```{r}
# Plot 
plot_cnv_area_padj <- function(df, tumour) {
  chr_levels <- c(as.character(1:22), "X", "Y")
  pd <- df %>%
    filter(CANCER_TYPE == tumour) %>%
    mutate(Arm        = str_extract(Cytoband, "^[0-9XY]+[pq]"),
           Chromosome = str_extract(Arm, "^[0-9XY]+"),
           Chromosome = factor(Chromosome, levels = chr_levels)) %>%
    arrange(Chromosome, Cytoband) %>%
    mutate(Cytoband = factor(Cytoband, levels = mixedsort(unique(Cytoband))),
           idx      = row_number())

  centromere_idx <- pd %>% filter(str_detect(Arm, "q")) %>%
    group_by(Chromosome) %>% slice_head(n = 1) %>% pull(idx)
  chr_last_idx <- pd %>% group_by(Chromosome) %>%
    slice_tail(n = 1) %>% pull(idx)
  chr_mid_idx <- pd %>% group_by(Chromosome) %>%
    summarise(idx = floor(mean(range(idx))), .groups = "drop")

  ggplot(pd, aes(idx, sign_logP, fill = Event)) +
    geom_area(position = "identity", alpha = 0.75) +
    geom_hline(yintercept = 0, linewidth = .25) +
   
     # Thresholds at padj < 0.2 and padj < 0.05
    geom_hline(yintercept = c(-log10(0.2), log10(0.2)),
               colour = "grey", linetype = "dotted", linewidth = .25) +
    geom_hline(yintercept = c(-log10(0.05), log10(0.05)),
               colour = "grey", linetype = "dotted", linewidth = .25) +
    geom_vline(xintercept = chr_last_idx + 0.5,
               colour = "grey50", linewidth = 0.3, linetype = "dashed") +
    geom_vline(xintercept = centromere_idx,
               colour = "grey80", linewidth = 0.2, linetype = "dashed") +
    scale_fill_manual(values = c("Gain" = "#E8437A", "Loss" = "#5BC4E5")) +
    scale_x_continuous(breaks = chr_mid_idx$idx,
                       labels = chr_mid_idx$Chromosome,
                       expand = c(0, 0)) +
    scale_y_continuous(limits = c(-1.3, 4), breaks = c(-1.3, 2, 4)) +
    labs(title = paste("Signed -log10(padj) CNV Survival –", tumour),
         x = "chromosome",
         y = "signed -log10(padj)",    # CHANGED: label reflects padj
         fill = "Event") +
    theme_bw() +
    theme(panel.border     = element_blank(),
          panel.grid       = element_blank(),
          axis.line        = element_line(linewidth = .25),
          axis.text.x      = element_text(size = 6),
          plot.title       = element_text(face = "bold", size = 12))
}

# ── STEP 3: Plot ──────────────────────────────────────────────────────────────
p_blca_padj <- plot_cnv_area_padj(results_cytoband_padj, "Bladder Cancer")
print(p_blca_padj)


ggsave("/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/MSK_survival/logHR_plots_updated/Bladder_Cancer_0.3_padj.pdf",
       plot = p_blca_padj, width = 10, height = 5, useDingbats = FALSE)
```


Furhter align to IS plot 
```{r}
# ── Load IS_CNV cytobands for alignment ───────────────────────────────────────
is_cnv_blca <- read.csv(
  "/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/TCGA/Resubmission_SCNAs/IS_CNV_cytolevel.csv",
  check.names = FALSE
)
is_cnv_cytobands <- unique(is_cnv_blca$Cytoband)
shared_bands <- intersect(cytoband_cols, is_cnv_cytobands)
cat("Shared bands:", length(shared_bands), "\n")
```

```{r}
# ── Fill in missing cytobands with 0 for alignment ───────────────────────────
all_combinations <- expand.grid(
  Cytoband   = shared_bands,
  Comparison = c("Gain", "Loss"),
  stringsAsFactors = FALSE
)

results_cytoband_aligned <- all_combinations %>%
  left_join(results_cytoband, by = c("Cytoband", "Comparison")) %>%
  mutate(
    CANCER_TYPE = "Bladder Cancer",
    HR          = ifelse(is.na(HR),       1, HR),
    log2HR      = ifelse(is.na(log2HR),   0, log2HR),
    sign_logP   = ifelse(is.na(sign_logP),0, sign_logP),
    P           = ifelse(is.na(P),        1, P),
    N           = ifelse(is.na(N),        0, N),
    Event       = ifelse(Comparison == "Gain", "Gain", "Loss")
  )

cat("Aligned results:", nrow(results_cytoband_aligned), "\n")
cat("Expected:", length(shared_bands) * 2, "\n")
```

```{r}
# Apply padj to aligned version
results_cytoband_aligned_padj <- results_cytoband_aligned %>%
  mutate(
    sign_logP = -log10(p.adjust(P, method = "BH")) * sign(log2HR)
  )

# run plot_cnv_area_padj
plot_cnv_area_padj <- function(df, tumour) {
  chr_levels <- c(as.character(1:22), "X", "Y")

  pd <- df %>%
    filter(CANCER_TYPE == tumour) %>%
    mutate(Arm        = str_extract(Cytoband, "^[0-9XY]+[pq]"),
           Chromosome = str_extract(Arm, "^[0-9XY]+"),
           Chromosome = factor(Chromosome, levels = chr_levels)) %>%
    arrange(Chromosome, Cytoband) %>%
    mutate(
      Cytoband = factor(Cytoband, levels = mixedsort(shared_bands)),
      idx      = as.integer(Cytoband)  # CHANGED: use factor level position, not row_number()
    )

  centromere_idx <- pd %>% filter(str_detect(Arm, "q")) %>%
    group_by(Chromosome) %>% slice_head(n = 1) %>% pull(idx)
  chr_last_idx <- pd %>% group_by(Chromosome) %>%
    slice_tail(n = 1) %>% pull(idx)
  chr_mid_idx <- pd %>% group_by(Chromosome) %>%
    summarise(idx = floor(mean(range(idx))), .groups = "drop")

  ggplot(pd, aes(idx, sign_logP, fill = Event)) +
    geom_area(position = "identity", alpha = 0.75) +
    geom_hline(yintercept = 0, linewidth = .25) +
    geom_hline(yintercept = c(-log10(0.2), log10(0.2)),
               colour = "grey", linetype = "dotted", linewidth = .25) +
    geom_hline(yintercept = c(-log10(0.05), log10(0.05)),
               colour = "grey", linetype = "dotted", linewidth = .25) +
    geom_vline(xintercept = chr_last_idx + 0.5,
               colour = "grey50", linewidth = 0.3, linetype = "dashed") +
    geom_vline(xintercept = centromere_idx,
               colour = "grey80", linewidth = 0.2, linetype = "dashed") +
    scale_fill_manual(values = c("Gain" = "#E8437A", "Loss" = "#5BC4E5")) +
    scale_x_continuous(breaks = chr_mid_idx$idx,
                       labels = chr_mid_idx$Chromosome,
                       expand = c(0, 0)) +
    scale_y_continuous(limits = c(-1.3, 4), breaks = c(-1.3, 2, 4)) +
    labs(title = paste("Signed -log10(padj) CNV Survival –", tumour),
         x = "chromosome",
         y = "signed -log10(padj)",
         fill = "Event") +
    theme_bw() +
    theme(panel.border = element_blank(),
          panel.grid   = element_blank(),
          axis.line    = element_line(linewidth = .25),
          axis.text.x  = element_text(size = 6),
          plot.title   = element_text(face = "bold", size = 12))
}
# plot 
p_blca_aligned_padj <- plot_cnv_area_padj(results_cytoband_aligned_padj, "Bladder Cancer")
print(p_blca_aligned_padj)

ggsave("/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/MSK_survival/logHR_plots_updated/Bladder_Cancer_aligned_padj.pdf",
       plot = p_blca_padj, width = 10, height = 5, useDingbats = FALSE)
```

## Include covariate 
```{r}
# ── Need to merge in TUMOR_PURITY and AneuploidyScore first ──────────────────
# Calculate aneuploidy score from cytoband data
aneuploidy_cytoband <- cnv_raw %>%
  select(SAMPLE_ID, all_of(cytoband_cols)) %>%
  mutate(AneuploidyScore = rowSums(abs(across(all_of(cytoband_cols))), na.rm = TRUE)) %>%
  select(SAMPLE_ID, AneuploidyScore)

# Add TUMOR_PURITY and AneuploidyScore to sample_info
sample_info <- clin_smp %>%
  select(SAMPLE_ID, PATIENT_ID, CANCER_TYPE, ONCOTREE_CODE, SAMPLE_TYPE, TUMOR_PURITY) %>%
  left_join(aneuploidy_cytoband, by = "SAMPLE_ID")

merged_sample <- sample_info %>%
  left_join(cnv_cat, by = "SAMPLE_ID")

df <- clin_pt %>%
  select(PATIENT_ID, OS_MONTHS, OS_STATUS, SEX, DRUG_TYPE) %>%
  left_join(merged_sample, by = "PATIENT_ID") %>%
  mutate(OS = if_else(str_detect(OS_STATUS, "LIVING"), 0, 1))
```

```{r}
# ── Run Cox regression WITH covariates (0.3 threshold) ────────────────────────
results_cytoband_adj <- data.frame()

tumor_data <- df %>% filter(CANCER_TYPE == "Bladder Cancer", DRUG_TYPE == "PD-1/PDL-1")

for (band in cytoband_cols) {
  for (comparison in c("Gain", "Loss")) {
    subset <- tumor_data %>%
      filter(!!sym(band) %in% c("None", comparison)) %>%
      mutate(CNV = factor(!!sym(band), levels = c("None", comparison))) %>%
      filter(!is.na(TUMOR_PURITY), !is.na(AneuploidyScore))

    if (nrow(subset) >= 10 && length(unique(subset$CNV)) == 2) {
      tryCatch({
        fit <- coxph(Surv(OS_MONTHS, OS) ~ CNV + TUMOR_PURITY + AneuploidyScore, data = subset)
        cnv_term <- grep("^CNV", names(coef(fit)), value = TRUE)
        if (length(cnv_term) != 1) next

        hr  <- exp(coef(fit)[cnv_term])
        p   <- summary(fit)$coefficients[cnv_term, 5]

        results_cytoband_adj <- rbind(results_cytoband_adj, data.frame(
          CANCER_TYPE = "Bladder Cancer",
          Cytoband    = band,
          Comparison  = comparison,
          HR          = round(hr, 3),
          log2HR      = log2(hr),
          P           = signif(p, 3),
          Event       = comparison,
          N           = nrow(subset)
        ))
      }, error = function(e) {
        message(paste("Skipped:", band, comparison, "-", e$message))
      })
    }
  }
}

results_cytoband_adj <- results_cytoband_adj %>%
  mutate(sign_logP = -log10(P) * sign(log2HR))

cat("Adjusted results:", nrow(results_cytoband_adj), "\n")

# Check 11p specifically
results_cytoband_adj %>% filter(str_detect(Cytoband, "^11p")) %>% arrange(P)
```

```{r}
# Plot
out_dir <- "/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/MSK_survival/logHR_plots_updated"
if (!dir.exists(out_dir)) dir.create(out_dir)

# ── Plot adjusted version using same function as colleague's code ─────────────
library(gtools)

plot_cnv_area_adj <- function(df, tumour, signif_cut = 1.3) {

  chr_levels <- c(as.character(1:22), "X", "Y")

  pd <- df %>%
    filter(CANCER_TYPE == tumour) %>%
    mutate(Arm        = str_extract(Cytoband, "^[0-9XY]+[pq]"),
           Chromosome = str_extract(Arm, "^[0-9XY]+"),
           Chromosome = factor(Chromosome, levels = chr_levels)) %>%
    arrange(Chromosome, Cytoband) %>%
    mutate(Cytoband = factor(Cytoband, levels = mixedsort(unique(Cytoband))),
           idx      = row_number())

  centromere_idx <- pd %>% filter(str_detect(Arm, "q")) %>%
    group_by(Chromosome) %>% slice_head(n = 1) %>% pull(idx)

  chr_last_idx <- pd %>% group_by(Chromosome) %>%
    slice_tail(n = 1) %>% pull(idx)

  chr_mid_idx <- pd %>% group_by(Chromosome) %>%
    summarise(idx = floor(mean(range(idx))), .groups = "drop")

  ggplot(pd, aes(idx, sign_logP, fill = Event)) +
    geom_area(position = "identity", alpha = 0.75) +
    geom_hline(yintercept = 0, linewidth = .25) +
    geom_hline(yintercept = c(-signif_cut, signif_cut),
               colour = "grey40", linetype = "dashed", linewidth = .25) +
    geom_vline(xintercept = chr_last_idx + 0.5, colour = "grey30", linetype = "dashed",linewidth = .3) +
    geom_vline(xintercept = centromere_idx, colour = "grey",
               linetype = "dashed", linewidth = .2) +
    scale_fill_manual(values = c("Gain" = "#E8437A", "Loss" = "#5BC4E5")) +
    scale_x_continuous(breaks = chr_mid_idx$idx,
                       labels = chr_mid_idx$Chromosome,
                       expand = c(0, 0)) +
    scale_y_continuous(limits = c(-2, 4), breaks = c(-2, 0, 2, 4)) +
    labs(title = paste("Signed -log10(p) CNV Survival (Adjusted for AS & purity) –", tumour),
         x = "chromosome",
         y = "signed -log(p)",
         fill = "Event") +
    theme_bw() +
    theme(panel.border     = element_blank(),
          panel.grid       = element_blank(),
          axis.line        = element_line(linewidth = .25),
          axis.text.x      = element_text(size = 6),
          plot.title       = element_text(face = "bold", size = 12))
}

# ── Print and save ────────────────────────────────────────────────────────────
p_blca_adj <- plot_cnv_area_adj(results_cytoband_adj, "Bladder Cancer")
print(p_blca_adj)

ggsave("/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/MSK_survival/logHR_plots_updated/BLCA_0.3_covariate.pdf",
       plot = p_blca_adj, width = 12, height = 4)

```

# 4 tumor types — aligned (padj), same method as survival_0.3_fixed
## Setup: 0.3 threshold, non-covariate Cox, keep every cytoband
```{r}
library(gtools)  # mixedsort()

# Load + categorize at 0.3 (same as BLCA)
clin_pt  <- read.delim("data_clinical_patient.txt", skip = 4, check.names = FALSE)
clin_smp <- read.delim("data_clinical_sample.txt",  skip = 4, check.names = FALSE)
cnv_raw  <- read.delim("all_cnv_table.txt", check.names = FALSE, quote = "\"") %>%
  tibble::rownames_to_column("SAMPLE_ID")

cytoband_cols <- grep("^[0-9XY]+[pq]", names(cnv_raw), value = TRUE)

cnv_cat <- cnv_raw %>%
  mutate(across(all_of(cytoband_cols), ~ case_when(
    .x >  0.3 ~ "Gain",
    .x < -0.3 ~ "Loss",
    TRUE      ~ "None"
  )))

df <- clin_pt %>%
  select(PATIENT_ID, OS_MONTHS, OS_STATUS, DRUG_TYPE) %>%
  left_join(
    clin_smp %>%
      select(SAMPLE_ID, PATIENT_ID, CANCER_TYPE, ONCOTREE_CODE) %>%
      left_join(cnv_cat, by = "SAMPLE_ID"),
    by = "PATIENT_ID"
  ) %>%
  mutate(OS = if_else(str_detect(OS_STATUS, "LIVING"), 0L, 1L))

# Optional: align band set to the IS plot (uncomment to use IS cytobands only)
# is_cnv       <- read.csv("/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/TCGA/Resubmission_SCNAs/IS_CNV_cytolevel.csv", check.names = FALSE)
# bands_to_use <- intersect(cytoband_cols, unique(is_cnv$Cytoband))
bands_to_use <- cytoband_cols            # default: keep every cytoband
cat("Bands used:", length(bands_to_use), "\n")
```

## Cox for all 4 tumors — ALWAYS keep a row (no missing cytobands)
```{r}
cancer_types_to_plot <- c(
  "Head and Neck Cancer", "Melanoma",
  "Non-Small Cell Lung Cancer", "Renal Cell Carcinoma"
)

results_4 <- data.frame()
for (tumor in cancer_types_to_plot) {
  td <- df %>% filter(CANCER_TYPE == tumor, DRUG_TYPE == "PD-1/PDL-1")
  cat("Running:", tumor, "- N:", nrow(td), "\n")
  for (band in bands_to_use) {
    for (comparison in c("Gain", "Loss")) {
      sub <- td %>%
        filter(!!sym(band) %in% c("None", comparison)) %>%
        mutate(CNV = factor(!!sym(band), levels = c("None", comparison)))
      hr <- NA_real_; p <- NA_real_; n <- nrow(sub)
      if (n >= 10 && length(unique(sub$CNV)) == 2) {
        tryCatch({
          fit <- coxph(Surv(OS_MONTHS, OS) ~ CNV, data = sub)
          hr  <- exp(coef(fit)[1]); p <- summary(fit)$coefficients[1, 5]
        }, error = function(e) {})
      }
      results_4 <- rbind(results_4, data.frame(   # <- row kept no matter what
        CANCER_TYPE = tumor, Cytoband = band, Comparison = comparison,
        Event = comparison,
        HR = ifelse(is.na(hr), 1, round(hr, 3)),
        log2HR = ifelse(is.na(hr), 0, log2(hr)),
        Raw_P_Cox = ifelse(is.na(p), 1, p), N = n, stringsAsFactors = FALSE
      ))
    }
  }
}
cat("Total rows:", nrow(results_4),
    " expected:", length(cancer_types_to_plot) * length(bands_to_use) * 2, "\n")
```

## Signed -log10(padj) + one shared genomic x-axis
```{r}
# BH-adjusted WITHIN each tumor. Same peak pattern as raw p, scaled down by FDR.
# HNSC/RCC keep strong peaks; NSCLC has 0 bands surviving FDR (honest result).
results_4 <- results_4 %>%
  group_by(CANCER_TYPE) %>%
  mutate(Padj      = p.adjust(Raw_P_Cox, method = "BH"),
         sign_logP = -log10(Padj) * sign(log2HR),
         sign_logP = ifelse(is.na(sign_logP) | is.infinite(sign_logP), 0, sign_logP)) %>%
  ungroup()

chr_levels <- c(as.character(1:22), "X", "Y")
lev        <- mixedsort(bands_to_use)          # single level vector -> identical axes

axis_ref <- tibble(Cytoband = factor(lev, levels = lev)) %>%
  mutate(idx = as.integer(Cytoband),
         Arm = str_extract(as.character(Cytoband), "^[0-9XY]+[pq]"),
         Chromosome = factor(str_extract(Arm, "^[0-9XY]+"), levels = chr_levels),
         arm_suffix = str_extract(Arm, "[pq]"))

centromere_idx <- axis_ref %>% filter(arm_suffix == "q") %>%
  group_by(Chromosome) %>% slice_min(idx, n = 1, with_ties = FALSE) %>% pull(idx)
chr_last_idx <- axis_ref %>% group_by(Chromosome) %>%
  slice_max(idx, n = 1, with_ties = FALSE) %>% pull(idx)
chr_mid_idx  <- axis_ref %>% group_by(Chromosome) %>%
  summarise(idx = floor(mean(range(idx))), .groups = "drop") %>% filter(!is.na(Chromosome))

# padj reference lines: -log10(0.05)=1.30, -log10(0.2)=0.70
sig_lines <- -log10(c(0.05, 0.2))
```

## Plot: individual aligned PDFs + one combined figure
```{r}
out_dir <- "logHR_plots_updated"
if (!dir.exists(out_dir)) dir.create(out_dir)

plot_tumor <- function(dat, tumour) {
  pd <- dat %>% filter(CANCER_TYPE == tumour) %>%
    mutate(Cytoband = factor(Cytoband, levels = lev), idx = as.integer(Cytoband)) %>%
    filter(!is.na(idx)) %>% arrange(idx)
  ggplot(pd, aes(idx, sign_logP, fill = Event)) +
    geom_area(position = "identity", alpha = .75) +
    geom_hline(yintercept = 0, linewidth = .25) +
    geom_hline(yintercept = c(-sig_lines, sig_lines),
               colour = "grey40", linetype = "dotted", linewidth = .25) +
    geom_vline(xintercept = chr_last_idx + 0.5, colour = "black", linewidth = .3) +
    geom_vline(xintercept = centromere_idx, colour = "grey30",
               linewidth = .25, linetype = "dashed") +
    scale_fill_manual(values = c(Gain = "#D64045", Loss = "#375E97")) +
    scale_x_continuous(breaks = chr_mid_idx$idx, labels = chr_mid_idx$Chromosome,
                       expand = c(0, 0)) +
    labs(title = paste("Signed -log10(padj) CNV Survival –", tumour),
         x = "chromosome", y = "signed -log10(padj)", fill = "Event") +
    theme_bw() +
    theme(panel.border = element_blank(), panel.grid = element_blank(),
          axis.line = element_line(linewidth = .25),
          axis.text.x = element_text(size = 6),
          plot.title = element_text(face = "bold", size = 12))
}

for (tm in cancer_types_to_plot) {
  ggsave(file.path(out_dir, paste0(gsub(" ", "_", tm), "_aligned_padj.pdf")),
         plot_tumor(results_4, tm), width = 10, height = 4, useDingbats = FALSE)
}

results_4f <- results_4 %>%
  mutate(Cytoband = factor(Cytoband, levels = lev), idx = as.integer(Cytoband),
         CANCER_TYPE = factor(CANCER_TYPE, levels = cancer_types_to_plot)) %>%
  filter(!is.na(idx))

p_all <- ggplot(results_4f, aes(idx, sign_logP, fill = Event)) +
  geom_area(position = "identity", alpha = .75) +
  geom_hline(yintercept = 0, linewidth = .25) +
  geom_hline(yintercept = c(-sig_lines, sig_lines),
             colour = "grey40", linetype = "dotted", linewidth = .25) +
  geom_vline(xintercept = chr_last_idx + 0.5, colour = "black", linewidth = .3) +
  geom_vline(xintercept = centromere_idx, colour = "grey30",
             linewidth = .25, linetype = "dashed") +
  scale_fill_manual(values = c(Gain = "#D64045", Loss = "#375E97")) +
  scale_x_continuous(breaks = chr_mid_idx$idx, labels = chr_mid_idx$Chromosome,
                     expand = c(0, 0)) +
  facet_wrap(~CANCER_TYPE, ncol = 1, strip.position = "right", scales = "free_y") +
  labs(title = "Signed -log10(padj) CNV Survival — 4 tumor types, aligned",
       x = "chromosome", y = "signed -log10(padj)", fill = "Event") +
  theme_bw() +
  theme(panel.grid = element_blank(), axis.text.x = element_text(size = 6),
        strip.text = element_text(face = "bold", size = 9),
        plot.title = element_text(face = "bold", size = 12))

ggsave(file.path(out_dir, "FourTumors_aligned_padj.pdf"),
       p_all, width = 11, height = 9, useDingbats = FALSE)
p_all

results_4 %>% group_by(CANCER_TYPE) %>%
  summarise(padj_lt_0.05 = sum(Padj < 0.05), padj_lt_0.2 = sum(Padj < 0.2),
            max_abs_signlogP = round(max(abs(sign_logP)), 2), .groups = "drop")
```
```

